Embedding Dimension of Polyhedral Losses

### 2020-01-30 colt-20/embed-dim.tex
Jessie reading through Bo's geometric proof

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}\label{sec:examples}
\bo{I suggest we don't use the notation $\elicembed$ at all and just write out $d$-embeddable every time. But don't feel strongly.}
\subsection{Abstain property, $\alpha \leq 1/2$}\label{subsec:example-abstain}
One classification-like problem that is of particular interest is the abstain property, elicited by the loss $\ellabs{1/2}$ given in Equation~\ref{eq:abstain}.
The property $\abstain{\alpha}$ can be verified:
\begin{equation}\label{eq:abstain-prop}
     \abstain{\alpha} = \begin{cases}
     \argmax_{y \in \Y} p(y) & \max_y p(y) \geq 1 - \alpha\\
     \bot & \text{otherwise}
     \end{cases}~.~
\end{equation}

\cite{ramaswamy2018consistent} study the abstain property in great length, presenting a $\ceil {\log_2(n)}$ dimension embedding of the abstain property.
However, it is unclear if this bound is tight, as the previously studied lower bounds of~\cite{ramaswamy2016convex} do not work well for this property, failing to give anything tighter than the trivial $1 \leq \elicembed(\abstain{\alpha})$ for $\alpha \leq 1/2$.

With our $1$-dimensional characterization, we already observe a tighter lower bound.
\begin{proposition}
	For $n \geq 3$ and $\alpha \leq 1/2$, the abstain loss $\ell_{\alpha}$ is not $1$-embeddable. \bo{In fact, for any $\alpha < 1$...}
\end{proposition}
\begin{proof}
	Consider the intersection graph of $\gamma := \abstain{\alpha}$: the node associated with $\gamma_\bot$ has $n$ edges, and since we assume $n \geq 3$, we fail to observe the requirement that every node of the intersection graph has degree at most $2$.
\end{proof}

@@

\begin{observation} \label{obs:qfp-invariant-lin}
  For any invertible $A \in \reals^{d \times d}$, if $V,\{X^j : j \in [\ell]\}$ is a feasible solution to the quadratic feasibility program, then so is $(VA),\{A^{-1}X^j : j \in [\ell]\}$.
\end{observation}
\begin{proof}
  The halfspace constraints are $(VA)(A^{-1}X^j) \leq B \iff VX^{j} \leq B$.
  The $j$th vertex constraint is a vector equality $\sum_{y \in \Y} p^j(y) (A^{-1}X^j)_y = \vec{0}$.
  If we let $a_m$ be the $m$th row of $A^{-1}$, then the $m$th row of the vector equality is
  \begin{align*}
    0 &= \sum_{y \in \Y} p^j(y) \inprod{a_m}{x^j_y}  \\
      &= \inprod{a_m}{\sum_{y \in \Y} p^j(y) x^j_y}  \\
      &= 0
  \end{align*}
  so the program is feasible.
\end{proof}

@@

\begin{lemma} \label{lemma:unique-spot-xiji}
  For $d=2$, any pair of linearly independent $v_i,v_j$ rule out all except for a unique set of points $x^{ij}_i$ and $x^{ij}_j$ from being feasible.
\end{lemma}
\begin{proof}
  From the halfspace constraints, we must have $\inprod{v_i}{x^{ij}_i} \leq -1$ and $\inprod{v_i}{x^{ij}_j} \leq 1$, which combines with Lemma \ref{lemma:xiji-minus-xijj} to give $\inprod{v_i}{x^{ij}_i} = 1$.
  This immediately also gives $\inprod{v_j}{x^{ij}_i} = -1$.
  This system of two inequalities in two dimensions has exactly one solution if $v_i,v_j$ are linearly independent.
\end{proof}

\begin{lemma} \label{lemma:180-degree-no-three}
  There is no feasible solution to the QFP for $\gamma_{\bot}$ and $\ell_{1/2}$ where three vectors $v_i,v_j,v_m$ lie strictly within a halfspace through the origin (i.e. all within $180^{\circ}$ of each other).
\end{lemma}
\begin{proof}
  Let three of the vectors be given, lying strictly inside a halfspace, and label them clockwise clockwise as $v_1,v_2,v_3$.
  WLOG suppose $v_1$ points vertically ``up'', as in Figure \ref{bo:todo}.
  By Lemma \ref{lemma:unique-spot-xiji}, the possible locations of the following points are all uniquely determined: $x^{ij}_i, x^{ij}_j$ for $(i,j) \in \{(1,2),(1,3),(2,3)\}$.
  Both points $x^{12}_1$ and $x^{13}_1$ lie on the line $\inprod{v_1}{x} = -1$, i.e. a horizontal line below the origin.
  We have constraints $\inprod{v_2}{x^{12}_1} = 1$ and $\inprod{v_2}{x^{13}_1} \leq 1$.
  This implies $x^{13}_1$ is left of $x^{12}_1$ on the line.
  But the symmetric constraints $\inprod{v_3}{x^{13}_1} \leq 1$ and $\inprod{v_3}{x^{12}_1} \leq 1$ imply that $x^{12}_1$ is left of $x^{13}_1$ on the line.
  This implies we must have $x^{12}_1 = x^{13}_1$.

@@

  This implies that the lines $\inprod{v_2}{x}=-1$ and $\inprod{v_3}{x}=1$, in particular, do not intersect anywhere within the bounds of $\inprod{v_1}{x} \in [-1,1]$.
  Therefore, either their intersection point $x^{23}_2$ or its negative $x^{23}_3$ violates the feasibility constraint $\inprod{v_1}{x} \leq 1$.
  This proves there is no feasible solution.
\end{proof}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}\label{sec:examples}
\bo{I suggest we don't use the notation $\elicembed$ at all and just write out $d$-embeddable every time. But don't feel strongly.}
\subsection{Abstain property, $\alpha \leq 1/2$}\label{subsec:example-abstain}
One classification-like problem that is of particular interest is the abstain property, elicited by the loss $\ellabs{\alpha}$ given in Equation~\ref{eq:abstain}.
The property $\abstain{\alpha}$ can be verified:
\begin{equation}\label{eq:abstain-prop}
     \abstain{\alpha} = \begin{cases}
     \argmax_{y \in \Y} p(y) & \max_y p(y) \geq 1 - \alpha\\
     \bot & \text{otherwise}
     \end{cases}~.~
\end{equation}

\cite{ramaswamy2018consistent} study the abstain property in great length, presenting a $\ceil {\log_2(n)}$ dimension embedding of the abstain property.
However, it is unclear if this bound is tight, as the previously studied lower bounds of~\cite{ramaswamy2016convex} do not work well for this property, failing to give any lower bound tighter than the trivial $1 \leq d$ for $\alpha \leq 1/2$, where $\abstain{\alpha}$ is $d$-embeddable.

With our $1$-dimensional characterization, we already observe a tighter lower bound.
\begin{proposition}
	For $n \geq 3$ and $\alpha \leq 1/2$, the abstain loss $\ell_{\alpha}$ is not $1$-embeddable. \bo{In fact, for any $\alpha < 1$...}
\end{proposition}
\begin{proof}
	Consider the intersection graph of $\gamma := \abstain{\alpha}$: the node associated with $\gamma_\bot$ has $n$ edges, and since we assume $n \geq 3$, we fail to observe the requirement that every node of the intersection graph has degree at most $2$.
	For an example with $n=3$, see Figure~\ref{fig:abstain-alpha-half}.
\end{proof}

We now use our $d$-dimensional characterization and some observations about the $\abstain{1/2}$ property to improve lower bounds from those given by~\cite{ramaswamy2016convex}.

@@

\begin{observation} \label{obs:qfp-invariant-lin}
  For any invertible $A \in \reals^{d \times d}$, if $V,\{X^j : j \in [\ell]\}$ is a feasible solution to the quadratic feasibility program, then so is $(VA),\{A^{-1}X^j : j \in [\ell]\}$.
\end{observation}
\begin{proof}
  The halfspace constraints are $(VA)(A^{-1}X^j) \leq B \iff VX^{j} \leq B$.
  The $j^{th}$ vertex constraint is a vector equality $\sum_{y \in \Y} p^j(y) (A^{-1}X^j)_y = \vec{0}$.
  If we let $a_m$ be the $m$th row of $A^{-1}$, then the $m^{th}$ row of the vector equality is
  \begin{align*}
    0 &= \sum_{y \in \Y} p^j(y) \inprod{a_m}{x^j_y}  \\
      &= \inprod{a_m}{\sum_{y \in \Y} p^j(y) x^j_y}  \\
      &= 0
  \end{align*}
  so the program is feasible.
\end{proof}

@@

\begin{lemma} \label{lemma:unique-spot-xiji}
  For $d=2$ and the level set $\gamma_\bot$ for $\ell_{1/2}$, any pair of linearly independent $v_i,v_j$ rule out all except for a unique set of points $x^{ij}_i$ and $x^{ij}_j$ from being feasible.
\end{lemma}
\begin{proof}
  From the halfspace constraints, we must have $\inprod{v_i}{x^{ij}_i} \leq -1$ and $\inprod{v_i}{x^{ij}_j} \leq 1$, which combines with Lemma \ref{lemma:xiji-minus-xijj} to give $\inprod{v_i}{x^{ij}_i} = 1$.
  This immediately also gives $\inprod{v_j}{x^{ij}_i} = -1$.
  This system of two inequalities in two dimensions has exactly one solution if $v_i,v_j$ are linearly independent.
\end{proof}

\begin{lemma} \label{lemma:180-degree-no-three}
  There is no feasible solution to the QFP for $\gamma_{\bot}$ and $\ell_{1/2}$ where three vectors $v_i,v_j,v_m$ lie strictly within a halfspace through the origin (i.e. all within $180^{\circ}$ of each other).
\end{lemma}
\begin{proof}
  Let three of the vectors be given, lying strictly inside a halfspace, and label them clockwise clockwise as $v_1,v_2,v_3$.
  WLOG suppose $v_1$ points vertically ``up'', as in Figure \ref{bo:todo}.
  By Lemma \ref{lemma:unique-spot-xiji}, the possible locations of the following points are all uniquely determined: $x^{ij}_i, x^{ij}_j$ for $(i,j) \in \{(1,2),(1,3),(2,3)\}$.
  Both points $x^{12}_1$ and $x^{13}_1$ lie on the line $\inprod{v_1}{x} = -1$, i.e. a horizontal line below the origin.
  We have constraints $\inprod{v_2}{x^{12}_1} = 1$ and $\inprod{v_2}{x^{13}_1} \leq 1$.
  This implies $x^{13}_1$ is left of $x^{12}_1$ on the line $\inprod{v_1}{x} = 1$. 
  But the symmetric constraints $\inprod{v_3}{x^{13}_1} \leq \jessie{=?} 1$ and $\inprod{v_3}{x^{12}_1} \leq 1$ imply that $x^{12}_1$ is left of $x^{13}_1$ on the line.
  This implies we must have $x^{12}_1 = x^{13}_1$.

@@

  This implies that the lines $\inprod{v_2}{x}=-1$ and $\inprod{v_3}{x}=1$, in particular, do not intersect anywhere within the bounds of $\inprod{v_1}{x} \in [-1,1]$.
  Therefore, either their intersection point $x^{23}_2$ or its negative $x^{23}_3$ violates the feasibility constraint $\inprod{v_1}{x} \leq 1$.
  This proves there is no feasible solution.
  \jessie{No feasible solution with three normals lying strictly in the same halfspace through the origin?}
\end{proof}
